---
title: "segmented"
output: html_document
---


## loading in of data

```{r}
xylem <- read.table("/Users/gturco/Documents/Projects/z_stack/data/experiments/all_d.txt", header=TRUE)
xylem_count <- read.table("/Users/gturco/Documents/Projects/z_stack/data/experiments/count_d.txt", header=TRUE)
xylem_count_logged <- read.table("/Users/gturco/Documents/Projects/z_stack/data/experiments/all_d_no_zero.txt", header=TRUE)


xylem <- as.data.frame(xylem)
xylem$VND7_log = log2(xylem$VND7)

xylem_count <- as.data.frame(xylem_count)
xylem_count$VND7_log = log2(xylem_count$VND7)


xylem_count_logged <- as.data.frame(xylem_count_logged)
xylem_count_logged$VND7_log = log2(xylem_count_logged$VND7)
xylem_count_logged$percent_xylem = log2(xylem_count_logged$percent_xylem)

```

## The R segemented pakage says it used a pieacewise linear regression to caluclate the breakpoints
## Both Mike and I agreeed that it looks like there are 4 breaks here rather than 2 and the fit could be better

```{r}
library(segmented)
library(ggplot2)

out.lm <- lm(percent_xylem ~ VND7_log, data = xylem)
o <- segmented(out.lm, seg.Z = ~VND7_log , psi = list(VND7_log = c(4,5)),
  control = seg.control(display = TRUE)
)

dat2 = data.frame(x = xylem$VND7_log, y = broken.line(o)$fit)

 p <- ggplot(data = xylem, aes(xylem$VND7_log, y = broken.line(o)$fit))+ geom_jitter(data = xylem, aes(xylem$VND7_log, y = xylem$percent_xylem),alpha = 0.7,  shape = 4)  
 
p + geom_line(data = dat2, color = 'blue')

davies.test(out.lm, seg.Z = ~VND7_log , k=4)

o

```



## Mike helped me write code on how to find breakpoints these are the breakpoints from my code 

```{r}
ggplot(data = xylem_count, aes(VND7_log, y=percent_xylem )) + geom_segment(y= 0, x= 0, xend = 5.312184, yend = 0, color="grey") +geom_segment(y= 85.29761, x= 6.585654, xend = 10, yend = 85.29761, color="grey") +  geom_segment(x= 5.312184, y= 0, xend = 6.585654, yend = 85.29761, color="grey") + geom_jitter(alpha = 0.7, size=0.2, shape = 4) + theme_bw() + theme(legend.position ="none", text=element_text(size=7), panel.margin = unit(0, "cm"), axis.ticks = element_line(size = 0.1), axis.line = element_line(size = 0.1), axis.ticks.length = unit(0.05, "cm"), axis.ticks.margin =unit(0, "cm")) + xlab(NULL) + ylab(NULL)

x <- sort(xylem$VND7_log)[29:50]
y <- xylem$percent_xylem[order(xylem$VND7_log)][29:50]

summary(lm(y~x))

## Slope
 lm(y ~ x)$coeff[2]
```

Here we see a much better fit to the data. This suggest that the slope is 16. Is there a way of measuring that 16 is a step slope? Or can we say our slope is 16 this is a steap slope suggesting a switch like function?

I still need to re-run this with the x-axis logged but need to re-go through all the images to do this. Since I did not include the unchanged xylem in my orginal anaysis just ectopic xylem formation. I need this information so both the count data and percentage do not start at zero and can be logged.



### All other code below is for future reseach and does not need to be disscused



## me playing around with segmented package

```{r}
library(segmented)
library(ggplot2)

set.seed(12)
xx <- 1:100
zz <- runif(100)
yy <- 2 + 1.5*pmax(xx - 35, 0) - 1.5*pmax(xx - 70, 0) + 15*pmax(zz - .5, 0) + 
  rnorm(100,0,2)
dati <- data.frame(x = xx, y = yy, z = zz)
out.lm <- lm(y ~ x, data = dati)
o <- segmented(out.lm, seg.Z = ~x, psi = list(x = c(30,60)),
  control = seg.control(display = TRUE)
)
dat2 = data.frame(x = xx, y = broken.line(o)$fit)

library(ggplot2)
ggplot(dati, aes(x = x, y = y)) +
  geom_point() +
  geom_line(data = dat2, color = 'blue')

broken.line(o)

davies.test(out.lm, seg.Z = ~x )
 pscore.test(out.lm, seg.Z = ~x)
 ##  β = 0 means no breakpoint.
 ## studies have shown that such Score test is more powerful than the Davies test (see reference) when
#the alternative hypothesis is ‘one changepoint’.
```

## loading in my datasets and figures code

```{r}
library(ggplot2)
xylem <- read.table("/Users/gturco/Documents/Projects/z_stack/data/experiments/all_d.txt", header=TRUE)
xylem_count <- read.table("/Users/gturco/Documents/Projects/z_stack/data/experiments/count_d.txt", header=TRUE)
xylem_count_logged <- read.table("/Users/gturco/Documents/Projects/z_stack/data/experiments/all_d_no_zero.txt", header=TRUE)


xylem <- as.data.frame(xylem)
xylem$VND7_log = log2(xylem$VND7)

xylem_count <- as.data.frame(xylem_count)
xylem_count$VND7_log = log2(xylem_count$VND7)


xylem_count_logged <- as.data.frame(xylem_count_logged)
xylem_count_logged$VND7_log = log2(xylem_count_logged$VND7)
xylem_count_logged$percent_xylem = log2(xylem_count_logged$percent_xylem)


ggplot(data = xylem, aes(VND7_log, y=percent_xylem ))+ geom_jitter(alpha = 0.5)    

ggplot(data = xylem, aes(VND7_log, y=percent_xylem )) + geom_segment(y= 0, x= 0, xend = 5.689121, yend = 0, color="grey") +geom_segment(y= 85.29761, x= 5.689121, xend = 10, yend = 85.29761, color="grey") +  geom_segment(x= 5.689121, y= 0, xend = 5.689121, yend = 85.29761, color="grey") + geom_jitter(alpha = 0.7, size=0.2, shape = 4) + theme_bw() + theme(legend.position ="none", text=element_text(size=7), panel.margin = unit(0, "cm"), axis.ticks = element_line(size = 0.1), axis.line = element_line(size = 0.1), axis.ticks.length = unit(0.05, "cm"), axis.ticks.margin =unit(0, "cm")) + xlab(NULL) + ylab(NULL)

#ggsave("/Users/gturco/Documents/Projects/xylem_R_figures/Figure_1/Switch.pdf", width=4.70, height=3.2, dpi=600, units="cm") 

ggplot(data = xylem, aes(VND7_log, y=percent_xylem )) + geom_segment(y= 0, x= 0, xend = 5.689121, yend = 0, color="grey") +geom_segment(y= 85.29761, x= 5.689121, xend = 10, yend = 85.29761, color="grey") +  geom_segment(x= 5.689121, y= 0, xend = 5.689121, yend = 85.29761, color="grey") + geom_jitter(alpha = 0.7, size=0.2, shape = 4) + facet_wrap(~ experiment) + theme_bw() + theme(legend.position ="none", text=element_text(size=7), panel.margin = unit(0, "cm"), axis.ticks = element_line(size = 0.1), axis.line = element_line(size = 0.1), axis.ticks.length = unit(0.05, "cm"), axis.ticks.margin =unit(0, "cm")) + xlab(NULL) + ylab(NULL)

#ggsave("/Users/gturco/Documents/Projects/xylem_R_figures/Figure_1/Switch_by_experment.pdf", width=4.70, height=3.2, dpi=600, units="cm") 


ggplot(data = xylem_count, aes(VND7_log, y=percent_xylem )) + geom_segment(y= 0, x= 0, xend = 5.689121, yend = 0, color="grey") +geom_segment(y= 85.29761, x= 5.689121, xend = 10, yend = 85.29761, color="grey") +  geom_segment(x= 5.689121, y= 0, xend = 5.689121, yend = 85.29761, color="grey") + geom_jitter(alpha = 0.7, size=0.2, shape = 4) + theme_bw() + theme(legend.position ="none", text=element_text(size=7), panel.margin = unit(0, "cm"), axis.ticks = element_line(size = 0.1), axis.line = element_line(size = 0.1), axis.ticks.length = unit(0.05, "cm"), axis.ticks.margin =unit(0, "cm")) + xlab(NULL) + ylab(NULL)

#ggsave("/Users/gturco/Documents/Projects/xylem_R_figures/Figure_1/Switch_count.pdf", width=4.70, height=3.2, dpi=600, units="cm") 


```





```{r}
ggplot(data = xylem_count, aes(VND7_log, y=percent_xylem )) + geom_segment(y= 0, x= 0, xend = 5.312184, yend = 0, color="grey") +geom_segment(y= 85.29761, x= 6.585654, xend = 10, yend = 85.29761, color="grey") +  geom_segment(x= 5.312184, y= 0, xend = 6.585654, yend = 85.29761, color="grey") + geom_jitter(alpha = 0.7, size=0.2, shape = 4) + theme_bw() + theme(legend.position ="none", text=element_text(size=7), panel.margin = unit(0, "cm"), axis.ticks = element_line(size = 0.1), axis.line = element_line(size = 0.1), axis.ticks.length = unit(0.05, "cm"), axis.ticks.margin =unit(0, "cm")) + xlab(NULL) + ylab(NULL)

x <- sort(xylem$VND7_log)[29:50]
y <- xylem$percent_xylem[order(xylem$VND7_log)][29:50]

summary(lm(y~x))


 lm(y ~ x)$coeff[2]
```

Look at standard deviation of errors in the middle and at the max and at the min - this would be an additional indicator of hysteresis
### Add bins for std


```{r}

lower_end <- subset(xylem, percent_xylem == 0)$VND7_log

mean(lower_end)
sd(lower_end)
#sem
sd(lower_end)/length(lower_end)


high_end <- subset(xylem, percent_xylem >= 75)$VND7_log

mean(high_end)
sd(high_end)
#sem
sd(high_end)/length(lower_end)


middle <- subset(xylem, percent_xylem < 75 & percent_xylem > 0 )$VND7_log

mean(middle)
sd(middle)
#sem
sd(middle)/length(middle)

```


```{r}

lower_end <- subset(xylem, VND7_log <= 4)$percent_xylem

mean(lower_end)
sd(lower_end)
#sem
sd(lower_end)/length(lower_end)


high_end <- subset(xylem, VND7_log >= 7)$percent_xylem

mean(high_end)
sd(high_end)
#sem
sd(high_end)/length(lower_end)


middle <- subset(xylem, VND7_log < 7 & VND7_log > 4 )$percent_xylem

mean(middle)
sd(middle)
#sem
sd(middle)/length(middle)

```

## Binning

```{r}

## to make the same order we see when ploting
x_axis <- xylem$VND7_log
y_axis <- xylem$percent_xylem
window_size <- 0.5
windows <- seq(min(x_axis),max(x_axis), by = window_size)

get_stats <- function(w, x_axis,y_axis, window_size) {
  cord <- which(x_axis >= w & x_axis <= w+ window_size)
  print(cord)
  se = sd(y_axis[cord])/length(y_axis[cord])
  return(se)
  #return(c(se,mean(y_axis[cord])))

}


y <- lapply(windows,get_stats, x_axis=x_axis, y_axis=y_axis,window_size=window_size )
SE <- unlist(y)
z <- data.frame(SE,windows)
z <- na.omit(z)

ggplot(z,aes(windows,SE)) + geom_point()
```


```{r}
x <- c(1:10, 13:22)
y <- numeric(20)
## Create first segment
y[1:10] <- 20:11 + rnorm(10, 0, 1.5)
## Create second segment
y[11:20] <- seq(11, 15, len=10) + rnorm(10, 0, 1.5)
## Plot it
par(mar=c(4,4,1,1)+0.2)
plot(x,y, ylim=c(5, 20), pch=16)

breaks <- x[which(x >= 9 & x <= 17)]

mse <- numeric(length(breaks))
for(i in 1:length(breaks)){
 piecewise1 <- lm(y ~ x*(x < breaks[i]) + x*(x>=breaks[i]))
 mse[i] <- summary(piecewise1)[6]
}
mse <- as.numeric(mse)
mse

breaks[which(mse==min(mse))]
piecewise2 <- lm(y ~ x*(x < 15) + x*(x > 15))
summary(piecewise2)
plot(x,y, ylim=c(5, 20), pch=16)


curve((6.7344 + 15.2354) + (0.5843-1.3025)*x, add=T, from=1, to=15)
curve((6.7344 - 1.2313) + 0.5843*x, add=T, from=15, to=max(x))
abline(v=13, lty=3)
```
```{r}

#Force a stright line to the begning and end of the data
### bin the data...
### Min - 0
### Max - 100
lower_end <- subset(xylem, VND7_log <= 4)
l <- lm(percent_xylem ~ VND7_log, data = lower_end)
summary(l)

middle <- subset(xylem, VND7_log < 7 & VND7_log > 4 )
m <- lm(percent_xylem ~ VND7_log, data = middle)
summary(m)


high <- subset(xylem, VND7_log >= 7 )
h <- lm(percent_xylem ~ VND7_log, data = high)
h <- lm(percent_xylem ~1, data = high)
summary(h)
plot(m)

high <- subset(xylem, VND7_log >= 8 )
intercept = mean(high$percent_xylem)
s <- mean(high$percent_xylem) - high$percent_xylem 
## SE mean for two stright lines...
## 3 compent system with lowest error
se <- sum(s^2)

91.64694
#y= intercept + slope/coeffent * xvalue
#y = intercept + 0
```
